---
title: "Full Tummy, Higher Cognitive Potential?"
author: "Izzul Fattah Aji Pratama"
output:
  html: 
---

::: panel-tabset
<!-- Task 1 Tab: Data and Documenting -->

## Data and Documenting

### **Downloading and Documenting Data Process**

1.  **Research Question**\
    My primary question is: is there a measurable relationship between a country’s level of undernourishment and its students’ cognitive performance based on PISA score?. I chose this because Indonesia’s new free lunch program for student (kindergarten to high school level) has ignited various debate (both praise for its potential and worries about its cost), and as an Indonesian citizen I’m keen to see whether the data bear out the assumption that better nutrition drives better learning. To unpack this I’ll look at two sub-questions:

-   Does undernourishment correlate differently with PISA Science, Math, and Reading scores?
-   How does Indonesia’s nutrition–PISA link compare to its Southeast Asian neighbors and the OECD average?

2.  **Data Sources**

-   FAO Undernourishment Inde**x** : provides an annual, country level estimate of the percentage of people whose dietary energy intake falls below minimum requirements.

-   OECD PISA Scores (Science, Math, Reading): measured every three years for 15 year old student in participating countries

    These datasets are ideal because they:

    -   Cover the same countries and overlapping years (2009–2022).

    -   Are publicly available and methodologically transparent.

    -   Offer standard, cross-national measures of nutrition sufficiency and educational outcomes.

3.  **Data Details**

-   **Data Type:**

    -   FAO Undernourishment Index: model-based estimate using national food balance sheets (administrative data) plus sample household surveys.

    -   PISA: stratified, probability-based sample survey of 15-year-old students

-   **Data Limitations:**

    -   For PISA data, Because it relies on stratified samples of 15-year-old student, some countries fail to meet minimum participation thresholds or there are no survey on a specific year. Those missing or incomplete data points appear as NA and can bias cross-country comparisons

    -   For FAO Undernourishment Index, the data is coming from modeled estimates. it only uses the average food supply and survey-based measures of intake variation on every households. If the underlying food or survey data are missing, delayed, or unreliable, those estimates can be unreliable and they can easily hide families who are truly going hungry.

-   Data Ethics & Privacy

    -   The data that being analyzed is publicly available under CC-BY 4.0 license and aggregate country level indicators. No individual or personally identifiable data is involved.

    -   Any findings should be presented without stigmatizing any country or community based on their cognitive scores or nutrition levels.

4.  **Data Acquisition and Processing Steps**
    -   **FAO Undernourishment CSV:**

        1.  Visit <https://data.worldbank.org/indicator/SN.ITK.DEFC.ZS>.

        2.  Click “Download CSV”

        3.  Save as `data/FAO Undernourishment Index.csv`.

    -   **PISA Excel:**

        1.  Go to <https://pisadataexplorer.oecd.org/ide/idepisa/>.

        2.  Select the “Reading” and “Mathematics”, “Science” from the "subject" tab and select the relevant countries and years to benchmark

        3.  Continue by clicking the "Grouping Variables" and select the student demographic. For this case, I select "all students" as the sample, not specific demographic profile

        4.  Continue by clicking "build the report" and download the data and save as `data/PISA Score.xls` with sheets `science_score`, `math_score`, `reading_score`.

    -   **Processing and Cleaning in R**

<!-- -->

1.  Load all the necessary function that will be used in the analysis

```{r, echo=TRUE, message=FALSE, warning=FALSE, results='hide'}
library(readxl)
library(tidyverse)
library(janitor)
library(stringr)
library(plotly)
```

2.  Read and Tidy the PISA data

    By default, the PISA data comes in `.xls` format and has multiple sheet in it called `sheet_1`, `sheet_2`, and `sheet_3` but I changed it manually to the related subject such as `math_score`, `science_score`, and `reading_score` to make it easier for processing in R.

    First thing to do is read the data using `read_excel` and skip 11 rows that contains metadata. After that, to make all variables are uniform in lower cases, I used `clean_names` . by default, since the data is in excel format, the author of the dataset was using merge cells when wrote the `year_study` in excel, so when I read it with R, it only show 4 observation of `year_study` and the rest of the row was NA values. to automatically fill the year value into any blank cell before there is a different year value that already filled. The process is run repeatedly.

    After that, I did restructuring the class of every variables and renamed it to make the analysis easier. Turns out there are still several variables that is not tidy yet. For example, there are several country name that has a suffix of year (e.g Albania (2015), Malaysia (2015), etc) that has a separate score value for that specific year. So I choose to integrate then into one similar row by removing the suffix using this `" \(\d{4}\)$"` syntax. All the process for tidying up the PISA data are the same for Math, Science and Reading data.

    After done with the tidying up process, I continue to plot the data using a line chart plot because I want to see the trend overtime and compared it to other countries. I choose several countries in Southeast Asian such as Brunei Darussalam", "Cambodia", "Indonesia", "Malaysia", "Philippines", "Singapore", "Thailand", "Viet Nam", and "International Average (OECD) to get a comparison. Furthermore, for the finishing part, I use `plotly` function so the chart could look more interactive.

    ```{r,echo=TRUE, message=FALSE, warning=FALSE, results='hide'}
    #| label: Read and Tidy PISA Math Data
    # Read and Tidy PISA Math Data

    pisa_math <- read_excel("data/PISA Score.xls", sheet = "math_score", skip=11) %>%
      clean_names() %>%
      fill(year_study) %>%
      mutate(
        year        = as.integer(year_study),
        country     = str_remove(jurisdiction, " \\(\\d{4}\\)$"),
        avg_math = as.numeric(average)
      ) %>%
      select(year, country, avg_math) %>%
      group_by(country, year) %>%                         
      summarise(
        avg_math = mean(avg_math, na.rm = TRUE),
        .groups  = "drop"
      ) %>%
      filter(!is.na(avg_math)) %>%   
      mutate(avg_math = na_if(avg_math, NaN))

    ```

    ```{r, echo=TRUE, message=FALSE, warning=FALSE, results='hide'}
    #| label: Read and Tidy PISA Reading Data
    # Read and Tidy PISA Reading Data

    pisa_reading <- read_excel("data/PISA Score.xls", sheet = "reading_score", skip=11) %>%
      clean_names() %>%
      fill(year_study) %>%
      mutate(
        year        = as.integer(year_study),
        country     = str_remove(jurisdiction, " \\(\\d{4}\\)$"),
        avg_reading = as.numeric(average)
      ) %>%
      select(year, country, avg_reading) %>%
      group_by(country, year) %>%                         
      summarise(
        avg_reading = mean(avg_reading, na.rm = TRUE),
        .groups  = "drop"
      ) %>%
      filter(!is.na(avg_reading)) %>%   
      mutate(avg_reading = na_if(avg_reading, NaN))


    ```

    ```{r, echo=TRUE, message=FALSE, warning=FALSE, results='hide'}
    #| label: Read and Tidy PISA Science Data
    # Read and Tidy PISA Science Data

    pisa_science <- read_excel("data/PISA Score.xls", sheet = "science_score", skip=11) %>%
      clean_names() %>%
      fill(year_study) %>%
      mutate(
        year        = as.integer(year_study),
        country     = str_remove(jurisdiction, " \\(\\d{4}\\)$"),
        avg_science = as.numeric(average)
      ) %>%
      select(year, country, avg_science) %>%
      group_by(country, year) %>%                         
      summarise(
        avg_science = mean(avg_science, na.rm = TRUE),
        .groups  = "drop"
      ) %>%
      filter(!is.na(avg_science)) %>%    
      mutate(avg_science = na_if(avg_science, NaN))

    ```

    ```{r, echo=TRUE, message=FALSE, warning=FALSE, results='hide'}
    #| label: PISA Science Plot
    # PISA Science Plot

    series <- c(
      "Brunei Darussalam", "Cambodia",    "Indonesia", 
      "Malaysia",         "Philippines", "Singapore", 
      "Thailand",         "Viet Nam",    "International Average (OECD)"
    )

    p <- pisa_science %>%
      filter(country %in% series) %>%
      ggplot(aes(x = year, y = avg_science, color = country)) +
        geom_line(size = 1) +
        geom_point(size = 2) +
        scale_x_continuous(
          breaks = c(2009, 2012, 2015, 2018, 2022),
          labels = c("2009", "2012", "2015", "2018", "2022")
        ) +
        labs(
          title    = "PISA Science Scores (2009–2022)",
          subtitle = "Southeast Asian Countries & OECD Average",
          x        = "Assessment Year",
          y        = "Average Science Score",
          color    = NULL            
        ) +
        theme_minimal(base_size = 14) +
        theme(
          legend.position   = "bottom",
          legend.text       = element_text(size = 10),
          legend.key.width  = unit(1.5, "lines"),
          axis.text.x       = element_text(hjust = 0.5),
          plot.margin       = margin(5, 5, 40, 5) 
        )

    ggplotly(p, tooltip = c("country","year","avg_science")) %>%
      layout(
        legend = list(
          orientation = "v",      
          x = 1.02,               
          y = 1,                  
          xanchor = "left",       
          yanchor = "top"        
        ),
        margin = list(l = 60, r = 120, t = 80, b = 50)
      ) %>%
      config(displayModeBar = FALSE)

    ```

    ```{r, echo=TRUE, message=FALSE, warning=FALSE, results='hide'}
    #| label: PISA Math Plot
    # PISA Math Plot

    m <- pisa_math %>%
      filter(country %in% series) %>%
      ggplot(aes(x = year, y = avg_math, color = country)) +
        geom_line(size = 1) +
        geom_point(size = 2) +
        scale_x_continuous(
          breaks = c(2009, 2012, 2015, 2018, 2022),
          labels = c("2009", "2012", "2015", "2018", "2022")
        ) +
        labs(
          title    = "PISA Math Scores (2009–2022)",
          subtitle = "Southeast Asian Countries & OECD Average",
          x        = "Assessment Year",
          y        = "Average Science Score",
          color    = NULL            
        ) +
        theme_minimal(base_size = 14) +
        theme(
          legend.position   = "bottom",
          legend.text       = element_text(size = 10),
          legend.key.width  = unit(1.5, "lines"),
          axis.text.x       = element_text(hjust = 0.5),
          plot.margin       = margin(5, 5, 40, 5)  
        )


    ggplotly(m, tooltip = c("country","year","avg_math")) %>%
      layout(
        legend = list(
          orientation = "v",      
          x = 1.02,               
          y = 1,                  
          xanchor = "left",      
          yanchor = "top"         
        ),
        margin = list(l = 60, r = 120, t = 80, b = 50)
      ) %>%
      config(displayModeBar = FALSE)

    ```

    ```{r echo=TRUE, message=FALSE, warning=FALSE, results='hide'}
    #| label: PISA Reading Plot
    # PISA Reading Plot

    r <- pisa_reading %>%
      filter(country %in% series) %>%
      ggplot(aes(x = year, y = avg_reading, color = country)) +
        geom_line(size = 1) +
        geom_point(size = 2) +
        scale_x_continuous(
          breaks = c(2009, 2012, 2015, 2018, 2022),
          labels = c("2009", "2012", "2015", "2018", "2022")
        ) +
        labs(
          title    = "PISA Reading Scores (2009–2022)",
          subtitle = "Southeast Asian Countries & OECD Average",
          x        = "Assessment Year",
          y        = "Average Science Score",
          color    = NULL       
        ) +
        theme_minimal(base_size = 14) +
        theme(
          legend.position   = "bottom",
          legend.text       = element_text(size = 10),
          legend.key.width  = unit(1.5, "lines"),
          axis.text.x       = element_text(hjust = 0.5),
          plot.margin       = margin(5, 5, 40, 5) 
        )

    ggplotly(r, tooltip = c("country","year","avg_reading")) %>%
      layout(
        legend = list(
          orientation = "v",      
          x = 1.02,               
          y = 1,                  
          xanchor = "left",      
          yanchor = "top"         
        ),
        margin = list(l = 60, r = 120, t = 80, b = 50)
      ) %>%
      config(displayModeBar = FALSE)

    ```

3.  Read and Tidy FAO data

    For the FAO data, first we read the data using `read_csv` and skip 11 row that contains the metadata. I also used `clean_names` to make all the variables name in lower case. Furthermore, I used a `pivot_long` function to transform two previous variables of `undernourishment_pct` and `year` into two separate tidy columns. I also remove the `indicator_code` that is only showing the code for `indicator_name` which is all the same.

    After done with the tidying up process, I continue to plot the data using a line chart plot because I want to see the trend overtime and compared it to other countries. I choose several countries in Southeast Asian such as Brunei Darussalam", "Cambodia", "Indonesia", "Malaysia", "Philippines", "Singapore", "Thailand", "Viet Nam", and "International Average (OECD) to get a comparison. Furthermore, for the finishing part, I use `plotly` function so the chart could look more interactive.

```{r, echo=TRUE, message=FALSE, warning=FALSE, results='hide'}
#| label: Read FAO data
# Read FAO Data

fao_data <- read_csv("data/FAO Undernourishment Index.csv", skip=4) %>%
  clean_names()

fao_clean <- fao_data %>%
  pivot_longer(
    cols          = starts_with("x"),     
    names_to      = "year",
    names_prefix  = "x",
    values_to     = "undernourishment_pct",
    names_transform = list(year = as.integer) 
  ) %>%
  filter(year >= 2009, year <= 2022) %>%
  select(-indicator_code)
```

```{r, echo=TRUE, message=FALSE, warning=FALSE, results='hide'}
#| label: Plot FAO Data
# Plot FAO Data

series_fao <- c(
  "Brunei Darussalam", "Cambodia",    "Indonesia", 
  "Malaysia",         "Philippines", "Singapore", 
  "Thailand",         "Viet Nam",    "OECD members")

f <- fao_clean %>%
  filter(country_name %in% series_fao) %>%
  ggplot(aes(x = year, y = undernourishment_pct, color = country_name)) +
    geom_line(size = 1) +
    geom_point(size = 2) +
    scale_x_continuous(
      breaks = c(2009, 2012, 2015, 2018, 2022)
    ) +
    labs(
      title    = "FAO Undernourishment Index (2009–2022)",
      subtitle = "Southeast Asian Countries & OECD Average",
      x        = "Assessment Year",
      y        = "Prevalence of Undernourishment (%)",
      color    = NULL
    ) +
    theme_minimal(base_size = 14) +
    theme(
      axis.text.x     = element_text(hjust = 0.5),
      axis.title.x    = element_text(margin = margin(t = 15)),
      plot.margin     = margin(5, 5, 5, 5)
    )

ggplotly(f, tooltip = c("country_name", "year", "undernourishment_pct")) %>%
  layout(
    legend = list(
      orientation = "v",
      x           = 1.02,
      y           = 1,
      xanchor     = "left",
      yanchor     = "top"
    ),
    margin = list(l = 60, r = 120, t = 80, b = 50)
  ) %>%
  config(displayModeBar = FALSE)
```

<!-- Task 2 Tab: Writing your blogpost -->

## Blog post

### Motivation

Earlier this year, a new Indonesia President elect, Prabowo Subianto, launched **“Makan Bergizi Gratis” (MBG)**, a free school lunch program for every student from kindergarten to senior high school level. The promise is simple, better nutrition should fuel brighter minds. Yet with such an enormous budget comes a bunch of questions. Will a daily healthy lunch actually boost students academic performance? Or would that budget be better spent boosting teacher pay, sharpening curriculum, or closing gaps in school facilities? As an Indonesian who’s watched both the excitement and the skepticism, I wanted to see what the data say, **is there a real link between a well fed classroom and higher cognitive results?**

### Data

To explore this, I combined two public datasets:

1.  Nutrition measure: FAO’s Prevalence of Undernourishment Index from country population (2009–2022). This gives a yearly, country-level snapshot of the share of people whose calorie intake falls below a healthy minimum around 2000 kcal/day.

2.  Cognitive measure: PISA (Programme for International Student Assessment) scores in Science, Math, and Reading (2009, 2012, 2015, 2018, 2022). PISA tests 15 year olds across hundreds of countries on real-world problem-solving skills. These sources are ideal because they cover the same countries and years, they’re methodologically transparent, and they let us compare Indonesia not only with its Southeast Asian neighbors but also with the International average.

### Analysis

```{r, echo=FALSE, message=FALSE, warning=FALSE}

#| label: Plot FAO Data

series_fao <- c(
  "Brunei Darussalam", "Cambodia",    "Indonesia", 
  "Malaysia",         "Philippines", "Singapore", 
  "Thailand",         "Viet Nam",    "OECD members")

f <- fao_clean %>%
  filter(country_name %in% series_fao) %>%
  ggplot(aes(x = year, y = undernourishment_pct, color = country_name)) +
    geom_line(size = 1) +
    geom_point(size = 2) +
    scale_x_continuous(
      breaks = c(2009, 2012, 2015, 2018, 2022)
    ) +
    labs(
      title    = "FAO Undernourishment Index (2009–2022)",
      subtitle = "Southeast Asian Countries & OECD Average",
      x        = "Assessment Year",
      y        = "Prevalence of Undernourishment (%)",
      color    = NULL
    ) +
    theme_minimal(base_size = 14) +
    theme(
      axis.text.x     = element_text(hjust = 0.5),
      axis.title.x    = element_text(margin = margin(t = 15)),
      plot.margin     = margin(5, 5, 5, 5)
    )

ggplotly(f, tooltip = c("country_name", "year", "undernourishment_pct")) %>%
  layout(
    legend = list(
      orientation = "v",
      x           = 1.02,
      y           = 1,
      xanchor     = "left",
      yanchor     = "top"
    ),
    margin = list(l = 60, r = 120, t = 80, b = 50)
  ) %>%
  config(displayModeBar = FALSE)

```

Before we dive into test scores, let’s check out Indonesia’s nutrition story. Back in 2009, Indonesia topped our Southeast Asia undernourishment chart, more people were missing their minimum calorie needs than anywhere else in the region. Over the next decade things got better, undernourishment steadily fell, hitting its low point around 2019. But then, starting in 2020, the trend reversed. More Indonesians slipped back into undernourishment, putting us again at the bottom of the regional rankings by 2022. You might think it caused by COVID-19 economic shocks, but our neighbors didn’t see the same setback. That makes you wonder if there’s more going on here than just a pandemic bump. But we wont criticize about government policy during those days :), let’s turn to the real question, how does all this tie into student performance on PISA?

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#| label: PISA Science Plot-Blog
# PISA Science Plot

series <- c(
  "Brunei Darussalam", "Cambodia",    "Indonesia", 
  "Malaysia",         "Philippines", "Singapore", 
  "Thailand",         "Viet Nam",    "International Average (OECD)"
)

p <- pisa_science %>%
  filter(country %in% series) %>%
  ggplot(aes(x = year, y = avg_science, color = country)) +
    geom_line(size = 1) +
    geom_point(size = 2) +
    scale_x_continuous(
      breaks = c(2009, 2012, 2015, 2018, 2022),
      labels = c("2009", "2012", "2015", "2018", "2022")
    ) +
    labs(
      title    = "PISA Science Scores (2009–2022)",
      subtitle = "Southeast Asian Countries & OECD Average",
      x        = "Assessment Year",
      y        = "Average Science Score",
      color    = NULL            
    ) +
    theme_minimal(base_size = 14) +
    theme(
      legend.position   = "bottom",
      legend.text       = element_text(size = 10),
      legend.key.width  = unit(1.5, "lines"),
      axis.text.x       = element_text(hjust = 0.5),
      plot.margin       = margin(5, 5, 40, 5) 
    )

ggplotly(p, tooltip = c("country","year","avg_science")) %>%
  layout(
    legend = list(
      orientation = "v",      
      x = 1.02,               
      y = 1,                  
      xanchor = "left",       
      yanchor = "top"        
    ),
    margin = list(l = 60, r = 120, t = 80, b = 50)
  ) %>%
  config(displayModeBar = FALSE)

```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#| label: PISA Math Plot-Blog
# PISA Math Plot

m <- pisa_math %>%
  filter(country %in% series) %>%
  ggplot(aes(x = year, y = avg_math, color = country)) +
    geom_line(size = 1) +
    geom_point(size = 2) +
    scale_x_continuous(
      breaks = c(2009, 2012, 2015, 2018, 2022),
      labels = c("2009", "2012", "2015", "2018", "2022")
    ) +
    labs(
      title    = "PISA Math Scores (2009–2022)",
      subtitle = "Southeast Asian Countries & OECD Average",
      x        = "Assessment Year",
      y        = "Average Science Score",
      color    = NULL            
    ) +
    theme_minimal(base_size = 14) +
    theme(
      legend.position   = "bottom",
      legend.text       = element_text(size = 10),
      legend.key.width  = unit(1.5, "lines"),
      axis.text.x       = element_text(hjust = 0.5),
      plot.margin       = margin(5, 5, 40, 5)  
    )


ggplotly(m, tooltip = c("country","year","avg_math")) %>%
  layout(
    legend = list(
      orientation = "v",      
      x = 1.02,               
      y = 1,                  
      xanchor = "left",      
      yanchor = "top"         
    ),
    margin = list(l = 60, r = 120, t = 80, b = 50)
  ) %>%
  config(displayModeBar = FALSE)


```

```{r echo=FALSE, message=FALSE, warning=FALSE}
#| label: PISA Reading Plot-Blog
# PISA Reading Plot

r <- pisa_reading %>%
  filter(country %in% series) %>%
  ggplot(aes(x = year, y = avg_reading, color = country)) +
    geom_line(size = 1) +
    geom_point(size = 2) +
    scale_x_continuous(
      breaks = c(2009, 2012, 2015, 2018, 2022),
      labels = c("2009", "2012", "2015", "2018", "2022")
    ) +
    labs(
      title    = "PISA Reading Scores (2009–2022)",
      subtitle = "Southeast Asian Countries & OECD Average",
      x        = "Assessment Year",
      y        = "Average Science Score",
      color    = NULL       
    ) +
    theme_minimal(base_size = 14) +
    theme(
      legend.position   = "bottom",
      legend.text       = element_text(size = 10),
      legend.key.width  = unit(1.5, "lines"),
      axis.text.x       = element_text(hjust = 0.5),
      plot.margin       = margin(5, 5, 40, 5) 
    )

ggplotly(r, tooltip = c("country","year","avg_reading")) %>%
  layout(
    legend = list(
      orientation = "v",      
      x = 1.02,               
      y = 1,                  
      xanchor = "left",      
      yanchor = "top"         
    ),
    margin = list(l = 60, r = 120, t = 80, b = 50)
  ) %>%
  config(displayModeBar = FALSE)

```

And voila! when we looked over to PISA, the picture isn’t much brighter. In 2022, Indonesia sits third from the bottom across Reading, Math, and Science, and the deepest decline since 2009 has come in Reading. Meanwhile our neighbors tell a different story. Singapore, for example, has climbed steadily to the top of the Southeast Asian rankings in every round of PISA in all aspects.

So, does better nutrition translate into higher test scores? If you plot undernourishment against PISA over time, there’s no simple one to one relationship. Some countries with very low hunger rates aren’t necessarily PISA standouts, and vice-versa, happened both in the Southeast Asian region or even if we compared to the International average. However that doesn’t mean providing nutritious meal is not important, but in the context of boosting academic performance, nutrition is not a single factor. It require a more holistic strategy such as investing in teacher training, upgrading school facilities, modernizing the curriculum, and supporting learning outside the classroom, too.

### Summary

I set out to see whether national undernourishment levels really have a direct correlation with student PISA scores and the short answer is: **weak correlation**. Using FAO’s undernourishment index and PISA Science, Math, and Reading results from 2009–2022, I find that, on average, countries with fewer undernourished people tend to score higher, but the pattern isn’t uniform across nations. Indonesia, for example, drove down hunger through 2019 yet saw no PISA gains especially in Reading. While neighbors in the region or even the average of international countries that could suppress the undernourished index doesn't necessarily has an increased on its score. In other words, better nutrition appears linked to stronger cognitive performance at a high level, but it’s far from a direct one to one relationship, meaning that alongside free school meals, comprehensive strategy in teaching quality, facilities, and curriculum are essential to truly boost learning.

### References

World Bank. (2025). *Prevalence of moderate or severe food insecurity (% of population) (Indicator SN.ITK.DEFC.ZS)*\[Data set\]. Retrieved June 13, 2025, from <https://data.worldbank.org/indicator/SN.ITK.DEFC.ZS>

Organisation for Economic Co-operation and Development. (n.d.). *PISA Data Explorer* \[Data set\]. Retrieved June 15, 2025, from <https://pisadataexplorer.oecd.org/ide/idepisa/>

```         
Wickham H, Averick M, Bryan J, Chang W, McGowan LD, François R, Grolemund G, Hayes A, Henry L, Hester J, Kuhn M,
  Pedersen TL, Miller E, Bache SM, Müller K, Ooms J, Robinson D, Seidel DP, Spinu V, Takahashi K, Vaughan D, Wilke
  C, Woo K, Yutani H (2019). “Welcome to the tidyverse.” _Journal of Open Source Software_, *4*(43), 1686.
  doi:10.21105/joss.01686 <https://doi.org/10.21105/joss.01686>.
```

```         
Wickham H, Bryan J (2025). _readxl: Read Excel Files_. doi:10.32614/CRAN.package.readxl
  <https://doi.org/10.32614/CRAN.package.readxl>, R package version 1.4.5,
  <https://CRAN.R-project.org/package=readxl>.
```

```         
Wickham H (2023). _stringr: Simple, Consistent Wrappers for Common String Operations_.
  doi:10.32614/CRAN.package.stringr <https://doi.org/10.32614/CRAN.package.stringr>, R package version 1.5.1,
  <https://CRAN.R-project.org/package=stringr>.
```

```         
Firke S (2024). _janitor: Simple Tools for Examining and Cleaning Dirty Data_. doi:10.32614/CRAN.package.janitor
  <https://doi.org/10.32614/CRAN.package.janitor>, R package version 2.2.1,
  <https://CRAN.R-project.org/package=janitor>.
```

```         
C. Sievert. Interactive Web-Based Data Visualization with R, plotly, and shiny. Chapman and Hall/CRC Florida,
  2020.
```

<!-- Task 3 Tab: Behind the Scenes -  -->

## Behind the Scenes

***Q1. Tell us about parts of your data processing or analysis that weren’t “sexy” and wouldn’t typically be included in a blog post. (e.g. Was their any data drudgery or time intensive wrangling? Were there any repetitive tasks or manual tasks? If it was easy, describe what made it easy?)***

Getting the FAO and PISA files themselves was quite smooth I would say, they come with good documentation and line up nicely by country and year. The headache started when I loaded the PISA Excel sheet: years were hidden in merged cells, some country names had “(2015)” or any random year tacked on, and everything read in as NAs until I spent time skipping rows, filling down the year column, stripping out those extra tags, and converting text to numbers. It wasn’t fun, but without that cleanup I couldn’t move forward.

***Q2. Were there any challenges that you faced in conducting this analysis. These may take the form of data limitations or coding challenges? (e.g. Was there anything in your analysis that you were not anticipating when you started? Did you have to change your intended scope? Did you need to master a new skill? Were there any problems you were proud of solving?)***

On the data side, I faced a challenge when I saw FAO didn’t list Singapore, even though Singapore is a PISA superstar. So I had to leave them out of that comparison and start comparing with other countries. On the coding side, tweaking the visuals took way longer than I expected. finding the right font sizes, margins, axis breaks, colors, and legend spots was all trial and error until it looked clear.

***Q3. Tell us about any imperfect parts of your work and how you would like to expand or improve this analysis in future? Be clear about any limitations or aspects of your analysis that fell beyond scope.***

I’m not totally happy with how the three PISA charts stack up. They take too much vertical space, and my gridExtra attempt to make a side by side chart also failed. Next time maybe I would try a faceted or patchwork layout to keep things compact and I would also like to bring another data related to education such as government spending to help explain the pattern more clearly.

***Q4. Also submit 4 earlier versions of your assignment to show your iterative process. These should be your messy versions and include exploratory code. We recommend you save these files as you progress through your assignment.***

here is my step by step process when working on this assignment :
:::

## 
